---
output: html_document
number_sections: yes
highlight: zenburn
authors: Meilin Moellenkamp and Nicolas Rosemberg
toc: yes
---


```{r, echo=FALSE, include=FALSE, error=FALSE}
### 1. Setting Working Directory & Loading the data
# Setting the Working Directory
# setwd("/Users/Meilin/Desktop/Collaborative Social Data/FinalProject")
setwd("/Users/Nico/Documents/Hertie/Social science data analysis/Final project/FinalProject")

# Loading the Citations
pkgs <- c("RJSONIO","WDI","dplyr","tidyr","knitr","DataCombine","lmtest","httr","XML","plyr","Amelia","XLConnect","countrycode","ggplot2","magrittr","fmsb","car","stargazer")
repmis::LoadandCite(pkgs, file = 'RpackageCitations.bib')

# Loading the merged files from the Working directory
Merged <- read.csv(file="MergedData")
# MergedOld <- read.csv(file="MergedOld")
```


```{r, echo=FALSE, include=FALSE, error=FALSE}

### 2. Load Required Packages

# install.packages("RJOSONIO")  
library(RJSONIO)
# install.packages("WDI")  
library(WDI)
# install.packages("dplyr")  
library(dplyr) 
# install.packages("tidyr")  
library(tidyr)
# install.packages("httr")  
library(httr) 
# install.packages("dplyr")  
library(dplyr)
# install.packages("XML")  
library(XML)
#install.packages("plyr")
library(plyr)
# install.packages("Amelia")  
library(Amelia) 
#install.packages("XLConnect")
library(XLConnect)    
# install.packages("countrycode")
library("countrycode")
# install.packages ("ggplot2")
library(ggplot2)
# # install.packages ("magrittr")
library(magrittr)
# install.packages ("fmsb")
library(fmsb)
# install.packages ("car")
library(car)
# install.packages("stargazer")
library(stargazer)

```

# WELCOME

Welcome to our site and thank you for your interest in our research project!

Our names are Meilin Moellenkamp and Nicolas Rosemberg and we are jointly conducting a study on the determinants of HIV/AIDS. 

# WHY ARE WE INTERESTED IN THIS TOPIC?  

That is a good question. We believe that understanding the determinants of HIV/AIDS is an important step before designing complex interventions that try to tackle this global problem. 

Is it really a global problem? While HIV/AIDS incidence rates are highly concentrated in Southern Africa, we do observe that most countries in our sample (most low- and middle-income countries) suffer from this disease:

**Global distribution of HIV/AIDS**

```{r, results='asis', echo=FALSE}
suppressPackageStartupMessages(library(googleVis))
IncidenceMap <- gvisGeoChart(Merged, locationvar = 'iso2c',
                        colorvar = 'Incidence',
                        options = list(
                          colors = "['#fff7bc', '#d95f0e']"
                        ))

print(IncidenceMap, tag = 'chart')
```

You are coming from one of the countries that are gray in the map and you might be wondering, *"Is HIV not a problem in my country?". Well, it might be! The gray countries are actually those countries for which we do not have sufficient data. 

# REDUCING HIV FIGHTS POVERTY

Besides making people healthier and happier, reducing HIV helps to fight poverty. Experts in the field of povery reduction designed a list of strategies to reduce global poverty and set a list of 8 goals in order to combat poverty. These goals are called the Millenium Development Goals and the deadline to reach this goal is only some months away (September 2015). For more information on the MDGs please refer to the following links:

*BACKGROUND INFORMATION*

**UNITED NATIONS**: http://www.un.org/millenniumgoals/

**ONE**: http://www.one.org/international/mdg/

*PROGRESS REPORT*

**CENTER FOR GLOBAL DEVELOPMEN*T*: http://www.cgdev.org/page/mdg-progress-index-gauging-country-level-achievements

Inspired by the puzzling question of why some countries failed to reach the HIV-related MDG we embarked this research.

# SO WHAT DID WE DO?


Thank you for visiting our website. If you have any questions, please contact us at m.moellenkamp@mpp.hertie-school.org or n.rosemberg@mpp.hertie-school.org